---
title: "04 Exercise"
author: "Your name here"
output:
  html_document:
    highlight: pygments
    theme: spacelab
  pdf_document: default
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

* * *

```{r load-packages, message=FALSE}
# Load the tidyverse package
library(tidyverse)
```

### 1: R variables and the assignment operator 
Assign the variable `my_var` the value 3
```{r ex1-assign}
my_var <- 3
```

We can see the value that's currently assigned to `my_var` by entering the name of the variable on its own line:

```{r ex1-display}
my_var
```

We can also use that value in later calculations:
```{r ex1-use}
my_var * 2
```

### 2. Reading data into R from spreadsheet files
```{r ex2-readdata}
library(readr)
nhanes <- read_csv("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/data/nhanes.csv")
```

### 3. Getting a first look at what's in the data frame
What command would you use to just show the first few rows of the data set? Try it! (Enter your ansewr in the code chunk below and run the entire report by hitting **Knit HTML**. Voila! The R output you need is already in your report.)
    
```{r ex3}
# enter your code for 3 here
```

### 4. Converting categorical variables to factors

```{r ex4}
# enter your code for 4 here
```

### 5. Listing distinct values of the Race variable

```{r ex5}
# enter your code here for listing the distinc values of the `Race` variable.
```

### 6. Summarizing Categorical Variables

```{r ex6}
# enter your code for Exercise 6 here
```
