---
title: "Week 2 Day 5"
subtitle: "Stat140-04"
author: "Shan Shan"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Learn R through a case study

Case study: birth weights and tobacco use during preganancy.

Are babies' birth weights affected by whether or not the mother used tobacco during pregnancy? Low birth weights are a risk factor for health problems later in life, so this is important!

Here is a data set with data on a sample of randomly selected babies who were born in North Carolina in 2004, with some information about the mother and the babies' weights in grams.

```{r, echo = FALSE, message = FALSE}
library(resampledata)
NCBirths2004 <- NCBirths2004[, 1:7]
```

```{r}
dim(NCBirths2004)
head(NCBirths2004)
```

Here's a plot of the data:

```{r, echo = FALSE}
ggplot(data = NCBirths2004, mapping = aes(x = Weight, color = Tobacco)) +
  geom_density()
```

#### Exercise 1

What statistics would you use to summarize the center and spread of the distribution of birth weights within each group?

#### Exercise 2: Calculating Summaries of Quantitative Variables: `summarize` and `group_by`

We saw from the plot that there seems to be a difference in birth weights. What are "typical" birth weights for the two groups?

```{r}
```

#### Exercise 3: Sorting the data with `arrange`

What was the shortest gestation time?

To answer this question, we will `arrange` the babies in increasing order of gestation time. Then, the babies with the shortest gestation times will be near the `head` of the data frame.

```{r}
NCBirths_by_gestation <- NCBirths2004 %>%
  arrange(Gestation)

head(NCBirths_by_gestation)
```

What was the longest gestation time?

Now we want to `arrange` the babies in `desc`ending order of gestation time, so the longest gestation times will be at the `head` of the data frame:

```{r}
```

#### Exercise 4: Modifying or Adding a Variable: `mutate`

The babies' birth `Weights` are currently recorded in grams. Let's add a new variable to the data frame called `Weight_lbs` with the babies' birth weights in pounds.

There are about 453.6 grams in a pound. We can divide the weight in grams by 453.6 to get the weight in pounds.

```{r}
```

As usual, we can take a look at the results by looking at the output from `head`.

```{r}
head(NCBirths2004)
```
