---
title: "Week 2 Day 5"
author: "Shan Shan"
subtitle: Stat140-04
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Learn R through a case study

Case study: birth weights and tobacco use during pregnancy.

Are babies' birth weights affected by whether or not the mother used tobacco during pregnancy? Low birth weights are a risk factor for health problems later in life, so this is important!

Here is a data set with data on a sample of randomly selected babies who were born in North Carolina in 2004, with some information about the mother and the babies' weights in grams.

```{r include=FALSE}
# you only need to do this once: after you finish downloading the fivethirtyeight package, you can remove this code chunk
# install.packages("resampledata")
```

```{r, echo = FALSE, message = FALSE}
library(resampledata)
NCBirths2004 <- NCBirths2004[, 1:7]
glimpse(NCBirths2004)
```

## Part I: Data visualization

#### Exercise 1

Make a boxplot for the babies' birth weights (`Weight`) in the dataset `NCBirths2004`.

```{r}

```

#### Exercise 2

Make a side-by-side boxplot to compare babies' birth weights (`Weight`) against whether the mother used Tobacco `Tobacco`

```{r}

```

#### Exercise 3

Make a histogram of the babies' birth weights colored by whether the mother used Tobacco. Choose an appropriate binwidth.

```{r}

```

#### Exercise 4

Make a density plot of the babies' birth weights colored by whether the mother used Tobacco

```{r}

```


#### Exercise 5

Make a density plot of the babies' birth weights **faceted** by whether the mother used Tobacco

```{r}

```


#### Exercise 6

Search on the internet how to

* add a title to your ggplot
* change the labels on the x and y axis
* change the legend titles

Modify your code in Exercise 4 so that

* x-axis is labeled as "Weight (g)"
* y-axis is labeled as "Number of babies"
* title is "Effect of Tobacco Use on Babies' Weight"
* legend title is "Tobacco Use"

```{r}

```

## Part II: Data summary

#### Exercise 1 

What statistics would you use to summarize the center and spread of the distribution of birth weights within each group?

#### Exercise 2: Calculating Summaries of Quantitative Variables: `summarize` and `group_by`

We saw from the plot that there seems to be a difference in birth weights. What are "typical" birth weights for the two groups?

```{r}

```

#### Exercise 3: Sorting the data with `arrange`

What was the lightest babies weight?

To answer this question, we will `arrange` the babies in increasing order of weight. Then, the babies with the lightest weight will be near the `head` of the data frame.

```{r}

```

What was the heaviest babies' weight?

Now we want to `arrange` the babies in `desc`ending order of gestation time, so the longest gestation times will be at the `head` of the data frame:

```{r}
```

#### Exercise 4: Modifying or Adding a Variable: `mutate`

The babies' birth `Weights` are currently recorded in grams. Let's add a new variable to the data frame called `Weight_lbs` with the babies' birth weights in pounds.

There are about 453.6 grams in a pound. We can divide the weight in grams by 453.6 to get the weight in pounds.

```{r}
```
