---
title: "Miniproject 1"
Author: "Type your name here"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: readable
editor_options: 
  markdown: 
    wrap: sentence
---

### 0. Load the tidyverse package

```{r global_options}
library(tidyverse)
```

### 1. R variables and the assignment operator

To get you started, run the following command by pressing the green arrow below.

```{r}
my_var <- 3
```

To do so, you can simply put your cursor on that line, and hit the **Run** button on the upper right corner of the pane.

This command instructs R to assign the variable `my_var` the value 3.
In Stat140 and in general statistical literacy, we use the word "variable" in two ways.
The first is a name that we've given a value that we want to be able to re-use later.
In this example, `my_var` is a variable.
We have *assigned* the value 3 to it using the *assignment operator*, `<-` (a less than sign followed by a minus sign, to form an arrow).

We can see the value that's currently assigned to `my_var` by entering the name of the variable on its own line:

```{r}
my_var
```

We can also use that value in later calculations:

```{r}
my_var * 2
```

Remember to run these code, you can simply put your cursor on that line, and hit the **Run** button on the upper right corner of the pane.

The second meaning of the word "variable" is more closely related to what we have discussed before: a column in a data frame.

### 2. Reading data into R from spreadsheet files

As you have seen in previous tutorials, the most common way to store data is in a data frame.
You can think of a data frame as being like a spreadsheet.
Each row corresponds to an observational unit, and each column corresponds to a variable.

Usually, the data are stored in a spreadsheet-like file outside of R.
The file format we'll work with most in this class is a csv file (csv stands for comma separated value).
We can read in csv files using the `read_csv` function, which is in the `readr` package:

The `nhanes` data is from the "National Health and Nutrition Examination Survey", a survey that is conducted every year and records demographic and health information about a large nationally-representative sample of Americans.
We will work with a relatively small subset of this data set; I have selected just a few variables and a subset of the participants from the 2011-2012 version of the survey.

```{r,  message=FALSE}
nhanes <- read_csv("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/data/nhanes.csv")
```

There are also functions to read in data from other file formats.
For example, if your data were stored in an excel file (with a file extension like xlsx), you could use the `read_excel` function from the `readxl` package to read the data in.
This function doesn't handle reading files from the internet very well yet, so we won't use it much in this class -- but it's there if you need it later.

### 3. Getting a first look at what's in the data frame

We've talked about three functions `head`, `str`, `dim`, `glimpse` in Exercise 02 that can be useful to get some rough ideas about the data.

What command would you use to just show the first few rows of the data set?
Try it!
(Enter your ansewr in the code chunk below and run the entire report by hitting **Knit HTML**. Voila! The R output you need is already in your report.)

```{r}
# enter your code for 3 here
```

### 4. Converting categorical variables to factors

When you first read a data set in, quantitative data types will usually be assigned the correct data type in R, but categorical variables will typically be stored as a character data type in R.
We'll need to tell R that these are categorical variables by converting them to `factors`.
A factor is just R's name for a categorical variable.

We will use the `mutate` function to modify the data frame.

```{r, echo = TRUE}
nhanes <- nhanes %>%
    mutate(Gender = factor(Gender))
```

The `%>%` operator is called the **piping** operator.
Basically, it takes the output of the current line and pipes it into the following line of code.
We will learn more about the piping operator next week.

**A note on piping:** Note that we can read these three lines of code as the following:

*"Take the `hnanes` dataset and **pipe** it into the `mutate` function. Using this mutate the variable `Gender` to make it into a categorical variable. Then assign this new resulting dataset to the object called `nhanes`, i.e. overwrite the old `nhanes` dataset with the new one containing the overwritten `Gender` variable."*

### 5. Listing distinct values of a variable

In order to know what to list for the possible levels of an ordinal categorical variable, you can use the `distinct` function to list the distinct values of the variable:

```{r, echo = TRUE}
nhanes %>% distinct(Education)
```

Now list the distinct values of the `Race` variable.
Write your code in the following code chunk.

```{r}
# enter your code here for listing the distinc values of the `Race` variable.
```

### 6. Summarizing Categorical Variables

It is often helpful to obtain counts of how many observational units fall into each category of a categorical variable, or into each combination of categories for two categorical variables.
We will do this with the table function:

```{r, echo = TRUE}
table(nhanes$Education, nhanes$Gender)
```

Use the `table()` functions to produce a contingency table for the variables `Education` and `Race`.

```{r}
# enter your code here for generating a contingency table for the variables `Race` variable.
```

## Resources for learning R and working in RStudio

These cheatsheets may come in handy throughout the semester:

-   [RMarkdown cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)
-   [Data transformation cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
-   [Data visualization cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)

Note that some of the code on these cheatsheets may be too advanced for this course.
However the majority of it will become useful throughout the semester.

## Lastly...

Once you finish, click on *Knit to PDF* and you'll see your document in a new pop-up window.
Name the file `MP01_[yourlastname].pdf` and submit the file on gradescope.
Each group only needs to submit once.
