---
title: "Week 7 Day 1"
subtitle: "Stat140-04"
output: 
  pdf_document: default
  html_document:
    highlight: pygments
    theme: cerulean
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, fig.align="center")
library(tidyverse)
library(knitr)
options(pillar.sigfig = 8)
```

# Wild Horses

What is the relationship between the size of a herd of horses and the number of foals (baby horses!!) that are born to that herd in a year?

```{r, echo = FALSE, message=FALSE}
horses <- read.csv("https://mhc-stat140-2017.github.io/data/sdm4/Wild_Horses.csv")
```

## 1. Plot the data

Use `ggplot()` to make a scatter plot with `Adults` being the explainatory variable and `Foals` being the response variable. Check the following conditions: (1) Is the trend approximately linear? (2) Are there obvious outliers? (3) Does the plot thicken?

## 2. Use the `lm()` function to fit a linear regression model to the data.

## 3. Use the model

Use 2 to write out the linear model. Interpret the slope and the intercept in context of the data and the model.

## 4. Look at histogram of residuals

Use the following code to check that the residuals follow a nearly normal distribution (and/or the sample size is large enough to apply the Central Limit Theorem)

```{r}
horses <- mutate(horses,
  residual = residuals(lm_fit),
  predicted = predict(lm_fit))
ggplot(data = horses, mapping = aes(x = residual)) +
  geom_density()
```

## 5. Inference and confidence interval

Based on 1 - 4. Should we do inference? If yes, give a 95% confidence interval for the slope coefficient. Is it significantly different than 0?

## 6. Hypothesis tests

Find the $p$p-value in the summary of your linear fit in 2. Based on this p-value, does the size of the heard appear to be a significant predictor of number of foals? Make sure to state the hypotheses and the significance level you are using.
