---
title: "Week 4 Day 3"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: spacelab
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
library(tidyverse)
```

### Introduction

There is only one *BIG* idea today, but this is an important one (perhaps the most important throughout this course if I dare to say).
That is

> *how the value of a sample statistic would vary from sample to sample, if random samples were randomly selected over and over from a population*

### Topic I: The blue M&M

I recently received a gift of 100,000 M&M's.
Since my favorite color of M&M's is blue (Guess why?), I would like to know what proportion of these M&M's are blue?

One way to answer this question would be to perform an exhaustive count.
However, this would be a long and tedious process.
Instead, let's do some sampling activity!
We are interested estimating the proportion of blue M&M's in a population size of 100,000 M&M's.

I have logged the information about these 100,000 M&M's I received in the following dataset.
Let's first load the data.

```{r, message = FALSE}
mm <- read_csv("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/u4d3/mm.csv")
```

#### Step 1: virtual sample once

Now let's take 50 random samples using the `sample_n` function.

```{r samp1, echo = TRUE}
samp1 <- mm %>%
  sample_n(50)
```

This command collects a simple random sample of size 50 from the `mm` dataset, and assigns the result to `samp1`.
This is similar to randomly drawing 50 M&M's from a non-transparent bag that contains the entire 100,000 M&M's.
Working with these 50 M&M's is considerably simpler than working with all 100,000 M&M's in the population.

Now compute the proportion of blue M&M's in `samp1` using the code below.

```{r samp1prop, echo = TRUE}
samp1 %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n))
```

What's the proportion of blue M&M's in `samp1`?

#### Step II: Help me please

Now your turn, run the following code and record the proportion of blue M&M's in your sample.

```{r samp_student, echo = TRUE}
samp_student <- mm %>%
  sample_n(50)

samp_student %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n))
```

How does the sample proportion of `samp_student` compare with that of `samp1`?

Not surprisingly, every time you take another random sample, you might get a different sample proportion.
The sample proportion is what we call a sample statistic.
The key idea I want you to get out of this exercise is

> *The value of a sample statistic would vary from sample to sample*

#### The key idea: sampling distribution

We saw that it is useful to get a sense of the distribution in the sample proportions, meaning *how the value of a sample statistic (i.e., proportion) would vary from sample to sample, if random samples were randomly selected over and over from a population*.
The distribution of sample proportions, termed as the *sampling distribution (of the proportion)*, is the key idea in this course.

The sampling distribution is useful, in our case, to estimate the true proportion of blue M&M's in the entire population.
The sampling distribution is centered at the true population proportion, and the spread of the distribution indicates how much variability is incurred by sampling only 50 M&M's at a time from the population.
Note that each time we computed a proportion, we did so from a **different** sample!

### Topic II: Hollywood movies

The dataset `HollywoodMovies` contains information on around 900 movies came out of Hollywood between 2007 and 2013.
You can find the data (ACS) at <http://www.lock5stat.com/datapage.html.>

Available variables:

+-----------------+--------------------------------------+
| Genre           | Action, Comedy, etc                  |
+-----------------+--------------------------------------+
| Budget          | Budget of making in thousand dollars |
+-----------------+--------------------------------------+
| Audience scores | Ratings from audience 0-100          |
+-----------------+--------------------------------------+

Use the StatKey website to answer the following questions.

#### Part I: Proportions of comedy films

a\.
What is the population and population parameter?

b\.
What is the sample and sample statistic?

c\.
When you click "Generate 1 sample," what is StatKey doing?

d\.
What did you notice in the sampling distribution (shape, center, spread) as we increase the sample size n?

#### Part II: Audience scores

a\.
What is the population and population parameter?

b\.
What is the sample and sample statistic?

c\.
What did you notice in the sampling distribution as we increase the sample size?
