---
title: "Unit 4 Tutorial: Exercise 13"
author: "Type your name here"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
library(tidyverse)
library(infer)
```

<!-- Download this following .Rmd template, copy the following code in the Rstudio console and hit enter: -->
<!-- ```{r klippy, echo=FALSE, include=TRUE} -->
<!-- klippy::klippy() -->
<!-- ``` -->

<!-- ```{r, include = TRUE, eval = FALSE} -->
<!-- download.file("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/u4d3/13Exercise.Rmd", destfile = "Ex13.Rmd") -->
<!-- ``` -->

<!-- Use the template to complete the following exercises. When you are done, submit the knitted pdf document on [Moodle](https://moodle.mtholyoke.edu/mod/assign/view.php?id=567542). -->


### Introduction

There is only one *BIG* idea today, but this is an important one (perhaps the most important throughout this course if I dare to say). That is 

> *how the value of a sample statistic would vary from sample to sample, if random samples were randomly selected over and over from a population*


### The blue M&M
I recently received a gift of 100,000 M&M's. Since my favorite color of M&M’s is blue (Guess why?), I would like to know what proportion of these M&M's are blue?

One way to answer this question would be to perform an exhaustive count. However, this would be a long and tedious process. Instead, let's do some sampling activity!
We are interested estimating the proportion of blue M&M's in a population size of 100,000 M&M's. 

I have logged the information about these 100,000 M&M's I received in the followoing dataset. Let's first load the data. 

```{r, echo = FALSE, message = FALSE}
mm <- read_csv("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/u4d3/mm.csv")
```

The name of the data frame is `mm` and the color variable is named as `mcolor`. Let's first take a look at the data by using the `glimpse` function


```{r}
glimpse(mm)
```

### Part I: virtual sample once
Now let's take 50 random samples using the `sample_n` function from the `infer` package. We will use the `infer` package often for the rest of the course.

```{r samp1, echo = TRUE}
samp1 <- mm %>%
  sample_n(50)

samp1
```

This command collects a simple random sample of size 50 from the `mm` dataset, and assigns the result to `samp1`. 
This is similar to randomly drawing 50 M&M's from a non-transparent bag that contains the entire 100,000 M&M's.
Working with these 50 M&M's is considerably simpler than working with all 100,000 M&M's in the population.

Now compute the proportion of blue M&M's in `samp1` using the code below. 
```{r samp1prop, echo = TRUE}
samp1 %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n))
```

What's the proportion of blue M&M's in `samp1`?

### Part II: Help me please

Now your turn, run the followoing code and record the proportion of blue M&M's in your sample on [this Google sheet](https://docs.google.com/spreadsheets/d/10H8E0POP1WWYg3XRJp6jI1Cc86vymnp6X7eBq993V3c/edit?usp=sharing).

```{r samp_student, echo = TRUE}
samp_student <- mm %>%
  sample_n(50)

samp_student %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n))
```

How does the sample proportion of `samp_student` compare with that of `samp1`? 

Not surprisingly, every time you take another random sample, you might get a different sample proportion. The sample proportion is what we call a sample statistic. The key idea I want you to get out of this exercise is

> *The value of a sample statistic would vary from sample to sample*


### Part III: Let's look at the numbers together
Let's plot the histogram of the $\hat{p}$ we collected from the class. 

```{r, eval= FALSE}
p_hat <- c()

sample_class <- data.frame(p_hat)

ggplot(data = sample_class, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Blue)",
    title = "Distribution of p_hat",
    subtitle = "Sample size = 50, Number of samples = 24 collected from everyone in class"
  )
```

### Part IV: Use R to automate this process
This is getting boring. Can we automate the process? 

Here, we use R to take () different samples of size 50 from the population, calculate the proportion of blue M&M's in each sample, and store each result in a vector called `sample_props50`. The function
`rep_sample_n` took a random sample of size `n` (50) from the population of all M&M's and repeat this sampling procedure `rep` () times. Note that we specify that replace = TRUE since sampling distributions are constructed by sampling with replacement.

```{r iterate, echo = TRUE}
sample_props50 <- mm %>%
  rep_sample_n(size = 50, reps = 24, replace = TRUE) %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n)) %>%
  filter(mcolor == "blue")

ggplot(data = sample_props50, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Blue)",
    title = "Distribution of p_hat",
    subtitle = "Sample size = 50, Number of samples = 24 collected from everyone in class"
  )
```


### Part V: Sample many many times
Why only sample so little times, Let's do more. Modify the followoing code block to sample 15000 times.

```{r iterate1, echo = TRUE}
sample_props50 <- mm %>%
  rep_sample_n(size = 50, reps = 24, replace = TRUE) %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n)) %>%
  filter(mcolor == "blue")

ggplot(data = sample_props50, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Blue)",
    title = "Distribution of p_hat",
    subtitle = "Sample size = 50, Number of samples = 24 collected from R"
  )
```

### The key idea: sampling distribution
From the previous questions, we saw that it is useful to get a sense of the distribution in the sample proportions, meaning *how the value of a sample statistic (i.e., proportion) would vary from sample to sample, if random samples were randomly selected over and over from a population*. The distribution of sample proportions, termed as the *sampling distribution (of the proportion)*, is the key idea in this course. 

The sampling distribution is useful, in our case, to estimate the true proportion of blue M&M’s in the entire population. The sampling distribution is centered at the true population proportion, and the spread of the distribution indicates how much variability is incurred by sampling only 50 M&M's at a time from the population. Note that for each of the 15,000 times we computed a proportion, we did so from a **different** sample!

To make sure you understand how sampling distributions are built, and exactly what the `rep_sample_n` function does, try modifying the code to create a sampling distribution of **25 sample proportions** from **samples of size 10**, and put them in a data frame named `sample_props_small`. Print the output. 

```{r iterate2}
sample_props_samll <- mm %>%
  rep_sample_n(size = 50, reps = 15000, replace = TRUE) %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n)) %>%
  filter(mcolor == "blue")
```

### Part VI: Effects of sample size

Explore the following questions with your group

**How does the sample size affect the sampling distribution?**

Use code chunks below to create sampling distributions of `p_hat` (proportion of blue M&M's) from samples of size 50, 100 and 1000. Use 15,000 simulations. 
How does the shape of the histogram (center and spread) change as the sample size increases? 

```{r}
set.seed(1)
sample_size <- 1

sample_props_sample_size <- mm %>%
  rep_sample_n(size = sample_size, reps = 15000, replace = TRUE) %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n)) %>%
  filter(mcolor == "blue")

sub <- paste0("Sample size = ", sample_size, ", Number of samples = 15000 collected from everyone in class")

ggplot(data = sample_props_sample_size, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.025) +
  labs(
    x = "p_hat (Blue)",
    title = "Distribution of p_hat",
    subtitle = sub
  )

```


**What if you increase the number of simulations?**

Modify the followoing code. Use 50, 100 and 1000 simulations. 
How does the shape of the histogram (center and spread) change as the number of simulations increases? 

```{r}
set.seed(1)
rep <- 1

sample_props_sample_size <- mm %>%
  rep_sample_n(size = 50, reps = rep, replace = TRUE) %>%
  count(mcolor) %>%
  mutate(p_hat = n /sum(n)) %>%
  filter(mcolor == "blue")

sub <- paste0("Sample size = 50,  Number of samples = ", rep, " collected from everyone in class")

ggplot(data = sample_props_sample_size, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Blue)",
    title = "Distribution of p_hat",
    subtitle = sub
  )

```