---
title: "Unit 4 Day 3"
output: 
  html_document:
    css: lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
library(klippy)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

### Sampling distribution

#### Learning objectives 
* Define sample statistic as a point estimate for a population parameter, for example, the sample mean is used to estimate the population mean, and note that point estimate and sample statistic are synonymous.

* Recognize that sample statistic (such as the sample mean) will vary from one sample to another, and define this variability as sampling distribution (sometimes also called sampling variation).

#### Before class
* Finish watching [sampling variability](https://www.coursera.org/learn/inferential-statistics-intro/lecture/lkQnZ/sampling-variability-and-clt) up to 4:30
* Complete `13 Quiz`: [VPN](http://138.110.84.108:3838/sshan/quiz/Unit4), [nonVPN](http://159.89.233.150:3838/stat140/quiz/Unit4)

#### In-class
* Slides posted after class
* Working on `13 Exercise`: please use the followoing code to download the exercise template --
```{r, include = TRUE, eval = FALSE}
download.file("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/u4d3/13Exercise.Rmd", destfile = "Ex13.Rmd")
```
* Optional: [openintro simulation](https://openintro.shinyapps.io/sampling_distributions/)

#### Items due
* `13 Quiz` due [on moodle](https://moodle.mtholyoke.edu/mod/assign/view.php?id=567540) 09/16 6pm EST
* `13 Exercise` due [on moddle](https://moodle.mtholyoke.edu/mod/assign/view.php?id=567543) 09/16 6pm EST