---
title: "Unit 4 Day 5"
output: 
  html_document:
    css: lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---
```{r setup, include=FALSE}
library(klippy)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

### Simulation-based inference II 

#### Learning objectives 
* Explain how the hypothesis testing framework resembles a court trial.
* Recognize that in hypothesis testing we evaluate two competing claims:
    * the null hypothesis, which represents a skeptical perspective or the status quo, and
    * the alternative hypothesis, which represents an alternative under consideration and is often represented by a range of possible parameter values.
* Construction of hypotheses:
    * Always construct hypotheses about population parameters (e.g. population mean, $\mu$) and not the sample statistics (e.g. sample mean, $\bar{x}$). 
    * Note that the population parameter is unknown while the sample statistic is measured using the observed data and hence there is no point in hypothesizing about it. 
    * Define the null value as the value the parameter is set to equal in the null hypothesis.
    * Note that the alternative hypothesis might be one-sided ($\mu <$ or $>$ the null value) or two-sided ($\mu \neq$ the null value), and the choice depends on the research question.
* Define a $p$-value as the conditional probability of obtaining a sample statistic at least as extreme as the one observed given that the null hypothesis is true. 
$$ p\text{-value} = P(\text{observed or more extreme sample statistic} \mid H_0 \text{true}) $$
* Infer that if a confidence interval does not contain the null value the null hypothesis should be rejected in favor of the alternative.

* Compare the $p$-value to the significance level to make a decision between the hypotheses:
    * If the $p$-value < the significance level, reject the null hypothesis since this means that obtaining a sample statistic at least as extreme as the observed data is extremely unlikely to happen just by chance, and conclude that the data provides evidence for the alternative hypothesis.
    * If the $p$-value > the significance level, fail to reject the null hypothesis since this means that obtaining a sample statistic at least as extreme as the observed data is quite likely to happen by chance, and conclude that the data does not provide evidence for the alternative hypothesis.
    * Note that we can never “accept” the null hypothesis since the hypothesis testing framework does not allow us to confirm it.
* Note that the conclusion of a hypothesis test might be erroneous regardless of the decision we make.
    * Define a Type 1 error as rejecting the null hypothesis when the null hypothesis is actually true.
    * Define a Type 2 error as failing to reject the null hypothesis when the alternative hypothesis is actually true.
* Note that the probability of making a Type 1 error is equivalent to the significance level, and choose a significance level depending on the risks associated with Type 1 and Type 2 errors.
    * Use a smaller α if Type 1 error is relatively riskier.
    * Use a larger α if Type 2 error is relatively riskier.

#### Before class
* Finish reading [Ingot example](https://sshanshans.github.io/stat140/reading/u4d5.html)

#### In-class
* Slides [pdf](http://sshanshans.github.io/stat140/slides/u4d5.pdf)
* Working on `15 Exercise`: please use the followoing code to download the exercise template --
```{r, include = TRUE, eval = FALSE}
download.file("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/u4d5/15Exercise.Rmd", destfile = "Ex15.Rmd")
```
* Introducing `EA 04`: Project proposal
  * [intro](https://sshanshans.github.io/stat140/eas/04/intro)
  * [grading guideline](https://docs.google.com/spreadsheets/d/1Wuk9FIaxYv3M-u7di6K13S_gnMhgbPA_Mqrtz7b3eso/edit?usp=sharing)
  * [moodle portal](https://moodle.mtholyoke.edu/mod/assign/view.php?id=568956)
  * peer evaluation form: link TBA

#### Items due
* `15 Exercise` due 09/17 6pm EST
* `EA 03` due 09/17 6pm EST
  * [intro](https://sshanshans.github.io/stat140/eas/03/intro)
  * [grading guideline](https://sshanshans.github.io/stat140/eas/03/grading.pdf)
  * [moodle portal](https://moodle.mtholyoke.edu/mod/assign/view.php?id=566087)
  * [peer evaluation form]( https://forms.gle/BKggBgTB1Bji7Liv8)