---
title: "Unit 5 Day 2"
output: 
  html_document:
    css: lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
library(klippy)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

###  Binomial distribution

#### Learning objectives 

* Determine if a random variable is binomial using the four conditions.
    * The trials are independent.
    * The number of trials, n, is fixed.
    * Each trial outcome can be classified as a success or failure.
    * The probability of a success, p, is the same for each trial.
* Calculate the number of possible scenarios for obtaining k successes in n trials using the `choose` function. 
* Calculate probability of a given number of successes in a given number of trials using the binomial distribution

#### Before class
* Finish watching [Introduction to binomial distribution](https://www.coursera.org/learn/probability-intro/lecture/kJ07f/binomial-distribution), [Example](https://www.coursera.org/learn/probability-intro/lecture/yef4b/working-with-the-binomial-distribution)
* Complete ```16 Quiz``` [VPN](http://138.110.84.108:3838/sshan/quiz/Unit5/), [nonVPN](http://159.89.233.150:3838/stat140/quiz/Unit5/)


#### In class
* Slides [pdf](http://sshanshans.github.io/stat140/slides/u5d2.pdf)
* Working on ```17 Exercise``` please use the followoing code to download the exercise template --
```{r, include = TRUE, eval = FALSE}
download.file("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/labs/u5d2/17Exercise.Rmd", destfile = "Ex17.Rmd")
```


#### Items due
* `16 Quiz` due [on moodle](https://moodle.mtholyoke.edu/mod/assign/view.php?id=570134) 09/22 6pm EST
* `17 Exercise` due [on moodle](https://moodle.mtholyoke.edu/mod/assign/view.php?id=570136) 09/22 6pm EST