---
title: "Unit 4 Tutorial: Exercise 14"
author: "Type your name here"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
library(tidyverse)
library(moderndive)
library(infer)
```


### The pennies activity
Let's first take a look at the pennies sample

```{r, echo = FALSE, message = FALSE}
glimpse(pennies_sample)
```

The name of the data frame is `pennies_sample` and the two variables are `ID` and `year`.

### Warm up
The following code chunk plots the histogram of the variable `year` and computes the mean of the variable `year`

```{r}
ggplot(pennies_sample, aes(x = year)) +
  geom_histogram(binwidth = 10, color = "white")

x_bar <- pennies_sample %>% 
  summarize(mean_year = mean(year))
x_bar
```

### Part I: virtual re-sample once
Now let's use the following code to perform the resampling with replacement of the 50 slips of paper representing our original sample 50 pennies:
```{r samp1, echo = TRUE}
virtual_resample <- pennies_sample %>% 
  rep_sample_n(size = 50, replace = TRUE)

virtual_resample
```

Observe how we explicitly set the `replace` argument to `TRUE` in order to tell `rep_sample_n()` that we would like to sample pennies with replacement.

Now let's compute the sample mean in this resample
```{r}
virtual_resample %>% 
  summarize(resample_mean = mean(year))
```

Record your `resample_mean` on [this Google sheet](https://docs.google.com/spreadsheets/d/10H8E0POP1WWYg3XRJp6jI1Cc86vymnp6X7eBq993V3c/edit?usp=sharing).

### Part II: Resample many times
Can we automate the process as before?

```{r iterate, echo = TRUE}
virtual_resamples <- pennies_sample %>% 
  rep_sample_n(size = 50, replace = TRUE, reps = 1)
virtual_resamples

virtual_resampled_means <- virtual_resamples %>% 
  group_by(replicate) %>% 
  summarize(mean_year = mean(year))
virtual_resampled_means

ggplot(virtual_resampled_means, aes(x = mean_year)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 1990) +
  labs(x = "Resample mean year")

```

### Part III: Constructing a confidence interval

```{r}
lci <- quantile(virtual_resampled_means$mean_year, 0.025)
rci <- quantile(virtual_resampled_means$mean_year, 0.975)

ggplot(virtual_resampled_means, aes(x = mean_year)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 1990) +
  shade_confidence_interval(endpoints = c(lci, rci)) +
  labs(x = "Resample mean year", caption = "Green lines represent 95% C.I. bounds")
```
