---
title: "Week 3 Day 2"
subtitle: "Stat140-04"
author: "Type your name here"
output:
  html_document:
    highlight: pygments
    theme: spacelab
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Fitting a linear model in R

### Part 1: Mortality and potion

#### Load data

Recall the mortality and hard water data from yesterday. We have recordings of the mortality rate (deaths per 100,000 population) and concentration of calcium in drinking water (parts per million) in 61 large towns in England and Wales

```{r load data, message = FALSE, warning = FALSE}
mortality_water <- read_csv("https://mhc-stat140-2017.github.io/data/sdm4/Hard_water_Derby.csv")
```

#### Problem 1 

Make a scatter plot with `Calcium` on the horizontal axis and `Mortality` on the vertical axis with the least square line

#### Problem 2 

Fit a linear model between the mortality rate `Mortality` and concentration of calcium `Calcium` using the `lm` function.

#### Problem 3

Find the slope, intercept in the output, and write out the linear equation.

#### Problem 4

One of the towns in our sample had a measured Calcium concentration of 71 parts per million. What is the predicted value for the mortality rate in that town?

#### Problem 5

Use the linear model to predict the mortality rate of a town with Calcium concentration of -3 parts per million. Does the predicted value make sense here?

## Part 2:

Biologists studying the effects of acid rain on wildlife collected data from 172 sites on streams in the Adirondack Mountains. Some of the sites are on the same stream. The researchers recorded the pH (acidity) of the water and the BCI, a measure of biological diversity. They are interested in determining if there is a relationship between water acidity and biological diversity. The following code reads the data in and takes a look at the first 10 rows of the data:

```{r}
streams <- read_csv("http://www.evanlray.com/data/sdm4/Streams.csv")
streams <- mutate(streams, BCI = as.numeric(BCI)) %>%
  filter(!is.na(pH) & !is.na(BCI)) %>%
  arrange(Stream)
head(streams, 10)

```

#### Problem 1

Which variable is the explanatory variable, and which is the response?

#### Problem 2

Make a scatter plot of the data with a linear regression overlaid on top.

#### Problem 3

Fit a linear model using the `lm` function.

#### Problem 4

Find the slope, intercept in the output, and write out the linear equation.

#### Problem 7

One of the site had a PH value of 7.1. What is the predicted value for the BCI in that site?

## Part 3: Anscombe's data

The following R code reads in a data set called `anscombe`. It has five pairs of $x$ variables and $y$ variables: $(x_1, y_1), (x_2, y_2), (x_3, y_3), (x_4, y_4),$ and $(x_5, y_5)$. 

#### Problem 1
Fit a linear model to each of the pairs, with the $y$ variable as the response and the $x$ variable as the explanatory variable. Print out a summary of the linear model fit. Take notes of the slope, intercept and $R^2$ for the linear model fit. What did you observe?

#### Problem 2 
Now make a scatter plot for each pair of x and y's. What did you observe? Is linear model useful here?

### Important message

Be sure to check the conditions for linear regression before reporting or interpreting a linear model.

From the scatterplot of y against x, check the

-   **Straight Enough Condition** Is the relationship between y and x straight enough to proceed with a linear regression model?

-   **Outlier Condition** Are there any outliers that might dramatically influence the fit of the least squares line?

-   **Does the Plot Thicken? Condition** Does the spread of the data around the generally straight relationship seem to be consistent for all values of x?
