---
title: "Week 3 Day 5"
author: "Type your name here"
subtitle: Stat140-04
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, highlight = TRUE)
anscombe <- read_csv("http://www.evanlray.com/data/base_r/anscombe_quintet.csv")

wildfires <- read_csv("http://www.evanlray.com/data/sdm4/Wildfires_2012.csv")
names(wildfires) <- c("num_fires", "years_since_1985", "ave_acres_burned")

wildfires <- wildfires %>%
  arrange(years_since_1985)
```

## I Anscombe's data

The following R code reads in a data set called `anscombe`. It has five pairs of $x$ variables and $y$ variables: $(x_1, y_1), (x_2, y_2), (x_3, y_3), (x_4, y_4),$ and $(x_5, y_5)$. You will examine one of these pairs of variables, and then we will discuss them all as a class. I'll tell you where the data came from later.

1.  Insert a code block below to fit a linear model to one of the pairs, with the $y$ variable as the response and the $x$ variable as the explanatory variable. Print out a summary of the linear model fit. Take notes of the slope, intercept and $R^2$ for the linear model fit. What did you observe?
2.  Now make a scatter plot for each pair of x and y's. What did you observe? Is linear model useful here?

### Important message

Be sure to check the conditions for linear regression before reporting or interpreting a linear model.

From the scatterplot of y against x, check the

-   **Straight Enough Condition** Is the relationship between y and x straight enough to proceed with a linear regression model?

-   **Outlier Condition** Are there any outliers that might dramatically influence the fit of the least squares line?

-   **Does the Plot Thicken? Condition** Does the spread of the data around the generally straight relationship seem to be consistent for all values of x?

## II Wildfire

The National Interagency Fire Center (www.nifc.gov) reports statistics about wildfires. I have loaded a dataset called `wildfires` with information about the number of wildfires, and the average number of acres burned per wildfire (in hundreds of thousands of acres) in each year from 1985 to 2012.

You can take a look at the data using the `head` function like this.

```{r}
head(wildfires)
```

### Part I

1.  Make a scatter plot with `years_since_1985` on the horizontal axis and `num_fires` on the vertical axis. Also add the least square line on your scatter plot.
2.  Fit a regression model using `num_fires` as the response variable and `years_since_1985` as the explanatory variable, and print out a summary of the linear model fit.
3.  Is a linear model appropriate for these data? For this we need to check the following conditions.
4.  Interpret the slope in this context.
5.  Can we interpret the intercept? Why or why not?
6.  What does $R^2$ mean in this context?

### Part II

We saw in Part I that the number of fires was nearly constant. But has the damage they cause remained constant as well? Let's examine the trend in Acres per Fire, (in hundreds of thousands of acres).

1.  Make a scatter plot with `years_since_1985` on the horizontal axis and `ave_acres_burned` on the vertical axis. Also add the least square line on your scatter plot.
2.  Fit a regression model using `ave_acres_burned` as the response variable and `years_since_1985` as the explanatory variable, and print out a summary of the linear model fit.
3.  Is a linear model appropriate for these data? Explain.
4.  Interpret the slope in this context.
5.  Can we interpret the intercept? Why or why not?
6.  What does $R^2$ mean in this context?

