---
title: "02 EA: Exploratory Data Analysis"
author: "INSERT YOUR NAME HERE"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: spacelab
---

```{r load-packages, message=FALSE, include= FALSE}
# Load packages
library(tidyverse) 
```


```{r load-data, message=FALSE, include = FALSE}
# Load data
college_recent_grads <- read_csv("https://raw.githubusercontent.com/sshanshans/stat140/master/docs/eas/02/recent-grads-140.csv")
```


### Exercise 1

(Type your answer to Exercise 1 here. Add code chunks as needed. Don't forget  to label your code chunk. Do not use spaces in code chunk labels.)

### Exercise 2

(Type your answer to Exercise 2 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

### Exercise 3

(Type your answer to Exercise 3 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

### Exercise 7

...

### Exercise 8

...

### Exercise 9

...

### Exercise 10

...

## Extra Credits

