---
title: "Week 5 Day 5"
author: "Stat140-04"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
library(tidyverse)
library(moderndive)
library(infer)
```

### The pennies activity

Let's first take a look at the pennies sample

```{r, echo = FALSE, message = FALSE}
glimpse(pennies_sample)
```

The name of the data frame is `pennies_sample` and the two variables are `ID` and `year`.

### Warm up

The following code chunk plots the histogram of the variable `year` and computes the median of the variable `year`

```{r}
ggplot(pennies_sample, aes(x = year)) +
  geom_histogram(binwidth = 10, color = "white")

x_bar <- pennies_sample %>% 
  summarize(median_year = median(year))
x_bar
```

### Part I: virtual re-sample once

Now let's use the following code to perform the resampling **with replacement** of the 50 slips of paper representing our original sample 50 pennies:

```{r samp1, echo = TRUE}
virtual_resample <- pennies_sample %>% 
  rep_sample_n(size = 50, replace = TRUE)

virtual_resample
```

Observe how we explicitly set the `replace` argument to `TRUE` in order to tell `rep_sample_n()` that we would like to sample pennies with replacement.

Now let's compute the sample mean in this resample

```{r}
virtual_resample %>% 
  summarize(resample_median = median(year))
```

### Part II: Resample many times

The following code creates 1000 bootstrap samples.

```{r iterate, echo = TRUE}
virtual_resamples <- pennies_sample %>% 
  rep_sample_n(size = 50, replace = TRUE, reps = 1000)

virtual_resampled_medians <- virtual_resamples %>% 
  group_by(replicate) %>% 
  summarize(median_year = median(year))

ggplot(virtual_resampled_medians, aes(x = median_year)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 1990) +
  labs(x = "Resample median year")

```

### Part III: Constructing a confidence interval

Let's first construct a confidence interval with the quantile method.

```{r}
lci <- quantile(virtual_resampled_medians$median_year, 0.025)
rci <- quantile(virtual_resampled_medians$median_year, 0.975)

ggplot(virtual_resampled_medians, aes(x = median_year)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 1990) +
  shade_confidence_interval(endpoints = c(lci, rci)) +
  labs(x = "Resample median year", caption = "Green lines represent 95% C.I. bounds from the quantile method")

lci
rci
```

Let's also construct a confidence interval with the SE method.

```{r}
#Compute the sample median from the original sample

pe <- median(pennies_sample$year)
se <- sd(virtual_resampled_medians$median_year)

lci_se <- pe - 2 * se
rci_se <- pe + 2 * se

ggplot(virtual_resampled_medians, aes(x = median_year)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 1990) +
  shade_confidence_interval(endpoints = c(lci_se, rci_se), color = "chocolate1", fill = "chocolate2") +
  labs(x = "Resample median year", caption = "Orange lines represent 95% C.I. bounds from the SE method")
```

Let's make an overlaid plot with the two confidence intervals. Note that you can change the displayed color with the color code provided here: <http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf>

```{r}
ggplot(virtual_resampled_medians, aes(x = median_year)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 1990) +
  shade_confidence_interval(endpoints = c(lci, rci)) +
  shade_confidence_interval(endpoints = c(lci_se, rci_se), color = "chocolate1", fill = "chocolate2") +
  labs(x = "Resample median year", caption = "Orange lines represent 95% C.I. bounds from the SE method, Green lines from the quantile method")

```

### On your own

The `moderndive` package also contains some interesting dataset. For example, the dataset `movies_sample`is a random sample of 32 action movies and 36 romance movies from <https://www.imdb.com/> and their ratings. We want to use this sample to understand what percentage of movies on IMDB are action movies.

```{r}
glimpse(movies_sample)
```

### Exercise 1

Create a bootstrap sample for `movies_sample`, and compute the proportion of Action movies in the resample.

### Exercise 2

Construct and visualize a bootstrap distribution with 1000 bootstrap samples for the proportion of Action movies on IMDB.

### Exercise 3

Determine the 95% bootstrap confidence interval based on the distribution you constructed above.

### Exercise 4

Interpret the confidence interval in context of the data.
