---
title: "Recap: Diamonds"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE)

```

```{r load-packages}
library(tidyverse)
library(DT) #to print nice tables
```

- What type of variable is price? 

Price is a numerical variable. 

- Would you expect its distribution to be symmetric, right-skewed, or left-skewed? Why? 

- Make a histogram of the distribution of diamond prices. Does the shape of the distribution match your expectation?

```{r}
ggplot(data=diamonds, mapping=aes(x=price)) + 
  geom_histogram() +
  labs(title = "Distribution of Diamond Prices")
```

The distribution of `price` is skewed right. 
<br>

- How many diamonds of each type of cut are there?

```{r}
diamonds %>%
  count(cut) 
```
<br>

- Calculate the relative frequency (proportion) of each clarity of diamonds.

```{r}
diamonds %>%
  count(clarity) %>%
  mutate(proportion = n/nrow(diamonds))
```
<br>

- Calculate the relative frequency of each clarity of diamonds by cut.

```{r}
diamonds %>%
  count(cut,clarity) %>%
  group_by(cut) %>%
  mutate(proportion = n /sum(n)) %>%
  datatable() #in the DT package
```
<br>

- Plot the relationship between depth and price of only fair cut diamonds.

```{r}
diamonds %>%
  filter(cut == "Fair") %>%
  ggplot(mapping=aes(x=depth, y=price)) +
  geom_point() +
  labs(title="Price vs. Depth", subtitle="For Fair Cut Diamonds")
```
<br>

- For each type of cut, calculate minimum (`min`), maximum (`max`), mean (`mean`), and median (`median`) price of diamonds of that type.

```{r}
diamonds %>%
  group_by(cut) %>%
  summarise(minimum = min(price), 
            maximum = max(price),
            mean = mean(price), 
            median = median(price))
```
