---
title: "Lab 00"
subtitle: "Logistics"
date: "01.10.19"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
library(knitr)
opts_chunk$set(fig.height = 2.65, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
	fig.align = "center",
	fig.height = 3.75,
	fig.width = 6.25,
	message = FALSE,
	warning = FALSE
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

## Meet Your TA!

---

## Introduce Yourself!

- Name

- Major or intended major

- Hobby / something you enjoy doing outside of school and work

---

## Today's Agenda

1. Research study survey

2. What to expect in lab

3. Questions about course policies

4. Getting to know you survey

5. Pretest

6. GitHub / RStudio Cloud

---

## Research study survey

- Elijah Carter: [benjamin.e.carter@duke.edu](mailto:benjamin.e.carter@duke.edu)

- Survey & Consent Form: [https://duke.qualtrics.com/jfe/form/SV_a3o7DGdWTELkcPr](https://duke.qualtrics.com/jfe/form/SV_a3o7DGdWTELkcPr)

---

## What to expect in lab

- Introduction to the lab assignment (~ 10 minutes)

- Work on the lab assignment (generally done in teams)

- All assignments will be completed using RStudio and submitted on GitHub (more on this next week)

- Today's lab will focus on some logistics to help prepare for the semester
---

## Any questions about course policies?

Course website: [http://bit.ly/sta199-sp19](http://bit.ly/sta199-sp19/)

---

## Getting to know you survey

1. Go to the course page on [sakai.duke.edu](https://sakai.duke.edu/). 

2. Find the Getting to know you survey under *Tests & Quizzes* 

3. Complete the entire survey. Brief answers are OK. 
<br><br>

Some of the information from this survey will be used to help with the formation of teams.

---

## Pretest

- Go to [http://bit.ly/sta199-sp19-pretest](http://bit.ly/sta199-sp19-pretest)

- Access Code: PLV7281FDJ

- Course: STA 199

- Section: 1

Do your best! Don't worry if you don't know the answers to the questions or are unsure. This information will only be used to assess the statistical background of students in the class and to help with the formation of teams.

---

## GitHub & RStudio Cloud

- If you do not have a GitHub account, go to [https://github.com/](https://github.com/) and create one. 
    - See [Happy Git with R](https://happygitwithr.com/github-acct.html#username-advice) for tips for selecting a username.
    
- If you had trouble joining RStudio Cloud or copying and completing the UN Votes analysis in class yesterday, try it again. Please ask your TA for help.

- If you were able to join RStudio Cloud and run the analysis, be sure to update your profile, so your name properly displays. Then you are free to leave.

