---
title: "Lab 02: Data wrangling and visualization"
author: ""
date: "01.24.19"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
library(knitr)
opts_chunk$set(fig.height = 2.65, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
# For nonsese...
library(emo)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

class: center, middle

# Welcome to lab

---

## Agenda

1. **Writing Exercise #1**

2. **Teams:** Find teammates.

3. **Lab introduction:** FiveThirtyEight article on college majors.

4. **Starting a new project:** Clone the repo into a new RStudio Cloud project.

5. **Configure git**: Make sure your git config is set up

6. **Warm-up:** Make sure you know how to push and pull.

---

## Writing Exercises

- There will be periodic writing exercises this semester to give you an opportunity to practice...
    - communicating statistical ideas in writing
    - giving and receiving constructive feedback to help refine your writing skills

- Each writing exercise will consist of three parts (about 15 minutes each): 
    - Respond to a prompt about a statistical concept and/or code 
    - Provide constructive feedback for classmates
    - Revise your response based on the peer feedback

- Writing exercises are 5% of the final course grade

---

## Writing Exercise 1

- The first writing exercise is about data visualization and wrangling. 
    - Response to the prompt - due tonight at 11:59p
    - Peer review - due Saturday at 11:59p
    - Final draft - due Monday at 11:59p

- You should wait to work on it after you've spent some time practicing data wrangling in today's lab. 

- If you finish lab early, you can go ahead and work on the first writing exercise. Otherwise, I will let you know when there are 15 minutes left in lab, so you have an opportunity to work on it before you leave.

---

class: middle, center

## Today's Lab 

---
    
## Teams

- Find your teammates: [STA 199 Teams](https://prodduke-my.sharepoint.com/:x:/g/personal/mt324_duke_edu/ES7_pDh-O0dKp4gGq2FitB8B3NAMO291NTf_PaatJ1_WxQ?e=9N9oMH)

- By the end of today's lab: 
    - Come up with a **team name** and let me know the name.
    - Find a **weekly 2-hour block between Thursday and Monday** that the team can meet to finish lab. You may not need to use the time every week, but it will be helpful to already have a plan for the weeks that you do need extra time.

---

## Lab introduction

- Lab instructions posted on the course website under <i class="fa fa-file-code-o fa-lg"></i>
    + Let's go find today's lab!
    
- [The Economic Guide To Picking A College Major](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/) by Ben Casselman
  
---

## Starting a new project

- Go to the course GitHub organization and find the **lab-02** repo that has your team name on it.

- On GitHub, click on the green Clone or download button, select Use HTTPS (this might already be selected by default, and if it is, you’ll see the text Clone with HTTPS). Click on the clipboard icon to copy the repo URL.

- Go to RStudio Cloud and into the course workspace. Create a New Project from Git Repo. You will need to click on the down arrow next to the New Project button to see this option.

- Copy and paste the URL of your assignment repo into the dialog box and click OK. 

---

## Configure Git

- Type the following lines of code in the console in RStudio filling in your name and email address.

```{r}
library(usethis)
use_git_config(user.name="your name", user.email="your email")
```

- Update the name of your RStudio project to "Lab 02 - Data wrangling and visualization"

- Before starting on lab, let's talk about the workflow now that you're on a team

---

## Team Workflow

- You and your teammates are working from the same lab-02 GitHub repo. This means you are all contributing to the same R Markdown document!

- For today's lab, only **one person** on the team should be typing the code/narrative in the R Markdown document at a time. 
    - Rotate who updates the document. Each team member should have at least one commit in the lab.
    - When you're not typing, you should still be contributing to the discussion but do not actually touch the files on your computer.

---

## Team Workflow

I will now demonstrate the steps for updating the lab documents and sharing the updated files with your teammates. 

- **Every team member**: Click the **Pull** button in the Git pane. This brings the most updated files from GitHub to your RStudio project. 

- **<u>ONE</u> team member**:
  - Knit the .Rmd file. 
  - Click the box next to <u>every</u> file in the Git pane.
  - Write a short and informative commit message. 
  - Push the changes to GitHub.

- **Every team member**: Click the **Pull** button in the Git pane. You should now see the changes reflected in the documents in your RStudio project.

---

## Warm-up

- Complete the warmup exercise to practice pushing and pulling changes.

- Once you've completed the warmup read through the tips on the next slide. You are then ready to start working on the rest of lab.
    - **Remember: only one person should updating the .Rmd file during the warmup**
    - We will talk more about working with teams in GitHub next week.

- Please raise your hand if you have any questions about the warmup exercise. 

---

## Tips

- Be ready to troubleshoot your document, since it will likely fail to knit on multiple occasions throughout the process. Read the error message carefully and take note of which line is preventing a successful knit.

- Make sure to keep track of your various chunks and to keep text and code in the right place.

- Remember that your R Markdown file is not aware of your project's global environment and can only make use of variables, functions, etc. that you have loaded or defined in the document.

- Remind yourself how the pipe operator (`%>%`) works.

- If you're unsure how a function works or what its arguments are, simply type `?` in front of it and hit enter (`?read_csv` for instance). The "Help" tab will open and provide a summary of the function as well as some examples.

---

## Writing Excerise 1

- Due **today, Thurs Jan 24 at 11:59p**

- You can find the prompt at [https://app.elireview.com](https://app.elireview.com)

- If you did not sign up for Eli review in class, follow the [instructions](https://www2.stat.duke.edu/courses/Spring19/sta199.001/slides/lec-slides/2b-tidy-data-wrangle.html#2) to sign up

- Please check your email on Friday for more instructions from Professor Tackett

