---
title: "Lab 09: Simulation-based inference"
author: ""
date: "03.28.19"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
library(knitr)
opts_chunk$set(fig.height = 2.35, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
# For nonsese...
library(emo)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```


## Agenda

- Next week midterm -- some tips and guidelines

- Lab 09 overview

- Writing Exercise 3 Peer Review (last 15 minutes of class)
    - [https://app.elireview.com](https://app.elireview.com)

---

## Midterm 02

- Assigned after class on Monday, due Friday at 11:59p

- On Monday: (at a minimum)
    - clone repo to RStudio
    - read the questions
    - make a plan for when you'll work on the exam
    
- Asking questions on Piazza and in office hours:
    - On Piazza - ask private questions to instructor and TA
    - Only ask questions to clarify exam instructions; We won't be reading/debugging code
    
- Exam content:
    - Mostly modeling and inference
    - Some exploratory data analysis (EDA)

---

### Lab 09: Simulation-based inference

**Data:** Random sample of births from North Carolina

**Goals:** 

- Understand risk factors for infant health
- Use simulation based testing to estimate and test

---

## Setting a seed

- We'll be randomly sampling, so make sure to set a seed.

- All members of a given team should use the same seed to get the same results.

---

## Lab flow

- Part 1: Baby weights

- Part 2: Baby weight vs. smoking

Each part requires some data wrangling and visualization along with the inference component

---

## Getting started with lab

.small[
- Go to the course GitHub organization and find the **lab-09** repo that has your team name on it.

- On GitHub, click on the green Clone or download button, select Use HTTPS. Click on the clipboard icon to copy the repo URL.

- Go to RStudio Cloud and into the STA199 course workspace. Create a New Project from Git repo. You will need to click on the down arrow next to the New Project button to see this option.

- Copy and paste the URL of your assignment repo into the dialog box and hit OK.

- Run the following in the console to configure Git: 

```{r eval=FALSE}
library(usethis)
use_git_config(user.name="your name", user.email="your email")
```
]