---
title: "Lab 10: Say cheers to web scraping!"
author: ""
date: "04.11.19"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta199-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
library(knitr)
opts_chunk$set(fig.height = 2.35, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
# For nonsese...
library(emo)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```


## Agenda

- Project reminder

- Lab 10 overview

---

## Project

- Proposals graded (check repo issues)

- Make sure you understand feedback

- Prioritize clarifying/perfecting your research question

---
### Lab 10: Cheers to web scraping!

**Data:** RateBeer.com: user reviews of beers and breweries

**Goals:** 

- Scrape North Carolina breweries page
- Scrape a state of your choice
- compare trends across these two states

**Use your lecture notes for help with code!**
---

## Getting started with lab

.small[
- Go to the course GitHub organization and find the **lab-10** repo that has your team name on it.

- On GitHub, click on the green Clone or download button, select Use HTTPS. Click on the clipboard icon to copy the repo URL.

- Go to RStudio Cloud and into the STA199 course workspace. Create a New Project from Git repo. You will need to click on the down arrow next to the New Project button to see this option.

- Copy and paste the URL of your assignment repo into the dialog box and hit OK.

- Run the following in the console to configure Git: 

```{r eval=FALSE}
library(usethis)
use_git_config(user.name="your name", user.email="your email")
```
]